Manuscript Figures

### 2021-05-31 src/manuscript_figures.Rmd
replace one-vs-all pancancer prediction with multi-class prediction - drop predict_labels function

```{r}
# Define a function that takes as input LFs and some patient labels (clinical/cancer type/subtype)
# and evaluates predictability of the labels using the LFs
predict_labels <- function(LFs, surv, tcga_subtypes, 
                           project, datatype, label.type, ...) {
  source(settings$utility_script)
  predict_stats <- do.call(rbind, sapply(simplify = F, names(LFs), function(tool) {
    require(data.table)
    message("predicting ",label.type," using ",tool)
    experiment <- paste(project, datatype, sep = '.')
    M <- LFs[[tool]][[experiment]]
    if(label.type == 'cancer_type') {
      labels <- data.frame('cancer_type' = surv[match(rownames(M), bcr_patient_barcode)]$project, 
                           row.names = rownames(M))
    } else if (label.type == 'subtype') {
      labels <- data.frame('subtype' = tcga_subtypes[match(rownames(M),
                                                           bcr_patient_barcode)]$Subtype_Selected, 
                           row.names = rownames(M))
    } else if (label.type == 'clinical') {
      labels <- process_clinical_covariates(rownames(M))
    }
    dt <- do.call(rbind, lapply(colnames(labels), function(x) {
      l <- labels[,x]
      if(is.numeric(l)){
        # categorize samples by intervals
        l <- as.character(cut_interval(l, 10))
      }
      l[is.na(l)] <- 'Undefined'
      message("Predicting ",x, " for experiment ",experiment)
      dt <- suppressMessages(predict_cluster_labels(M, l, ...))
      dt$label.type <- x
      dt$project <- project
      dt$datatypes <- datatype
      dt$method <- tool
      return(dt)
    }))
    return(dt)
  }))
  return(predict_stats[!is.na(train_auc)])
}

plot_comp1 <- function(dt) {
  dt$method <- toupper(gsub("_factors", "", dt$method))
  comps <- lapply(setdiff(unique(dt$method), 'MAUI'), function(x) c('MAUI', x))
  ggboxplot(dt, 
          x = 'method', y = 'test_auc', color = 'method', 
          add = 'jitter', legend = 'none') + 
    stat_compare_means(method = 'wilcox.test', paired = T, 
                       method.args = list('alternative' = 'greater'), 
                     comparisons = comps) +
    labs(y = 'AUC') + theme(axis.text = element_text(size = 10))
}

@@

```{r pancancer_prediction, fig.height=9, fig.width=12}
# Find out if cancer types are predictable using latent factors 
pancancer_stats <- predict_labels(LFs, surv, tcga_subtypes, 'pancancer', 
                                  'cnv_gex_meth_mut', label.type = 'cancer_type', num.trees = 1000)
p1 <- plot_comp1(pancancer_stats) + theme(axis.title.x = element_blank())
p2 <- plot_comp2(pancancer_stats)
# plot cancer specific factors
M <- LFs$maui_factors$pancancer.cnv_gex_meth_mut
ct <- surv[match(rownames(M), bcr_patient_barcode)]$project
p3 <- plot_label_specific_vars(M, ct)

>>>>

```{r}
plot_comp1 <- function(dt) {
  dt$method <- toupper(gsub("_factors", "", dt$method))
  comps <- lapply(setdiff(unique(dt$method), 'MAUI'), function(x) c('MAUI', x))
  ggboxplot(dt, 
          x = 'method', y = 'test_auc', color = 'method', 
          add = 'jitter', legend = 'none') + 
    stat_compare_means(method = 'wilcox.test', paired = T, 
                       method.args = list('alternative' = 'greater'), 
                     comparisons = comps) +
    labs(y = 'AUC') + theme(axis.text = element_text(size = 10))
}

@@

```{r pancancer_prediction, fig.height=9, fig.width=12}
# Find out if cancer types are predictable using latent factors

res <- sapply(simplify = F, names(LFs), function(tool) {
  message(date(), " => ", tool)
  M <- LFs[[tool]]$pancancer.cnv_gex_meth_mut
  ct <- gsub("-", "_", surv[match(rownames(M), bcr_patient_barcode)]$project)
  res <- predict_labels_glmnet_multiclass(M, ct, partition = 0.6, nodes = 20)
})

pancancer_stats <- sapply(res, function(x) {
  caret::multiClassSummary(x, lev = levels(x$obs))
})

dt <- data.table(melt(pancancer_stats))[Var1 %in% c('Mean_Balanced_Accuracy')]
dt$Var2 <- toupper(gsub("_factors", "", dt$Var2))
p1 <- ggplot(dt, aes(x = Var2, y = value)) + 
  geom_bar(stat = 'identity', aes(fill = Var2)) + 
  labs(y = 'Mean Balanced Accuracy') + 
  theme(legend.title = element_blank(), axis.title.x = element_blank())

# performance visualisation
m <- as.matrix(table(res$maui_factors$pred, res$maui_factors$obs))
p2 <- pheatmap(apply(m, 1, function(x) round(x/sum(x) * 100, 1)), cluster_rows = F, cluster_cols = F, 
         display_numbers = T, number_format = "%.1f", silent = T)

# plot cancer specific factors
M <- LFs$maui_factors$pancancer.cnv_gex_meth_mut
ct <- surv[match(rownames(M), bcr_patient_barcode)]$project
p3 <- plot_label_specific_vars(M, ct)
